'White Wine Quality' Data Analysis using R by Ng Fang Kiang
========================================================
White Wine Quality: Click [Here](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityWhites.csv) to download.
Sources: Created by: Paulo Cortez (Univ. Minho), Antonio Cerdeira, Fernando Almeida, Telmo Matos and Jose Reis (CVRVV) @ 2009

This dataset contains 4,898 rows of white wines records data and 11 columns of variables on quantifying the chemical properties of each wine. The last columns of this dataset have recorded the output of this dataset which is the quality of the wine that rated by at least 3 wine experts. The ratings are rated from the lowest of 0 to 10 that represented "VERY POOR" to "VERY EXCELLENT"!

## Description of attributes:

### Input variables (based on physicochemical tests):
1. **fixed.acidity** (tartaric acid - g / dm^3): most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
2. **volatile.acidity** (acetic acid - g / dm^3): volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
3. **citric.acid** (g / dm^3): found in small quantities, citric acid can add 'freshness' and flavor to wines
4. **residual.sugar** (g / dm^3): the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet
5. **chlorides** (sodium chloride - g / dm^3): the amount of salt in the wine
6. **free.sulfur.dioxide** (mg / dm^3): the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine
7. **total.sulfur.dioxide** (mg / dm^3): amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
8. **density** (g / cm^3): the density of water is close to that of water depending on the percent alcohol and sugar content
9. **pH**: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
10. **sulphates** (potassium sulphate - g / dm3): a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant
11. **alcohol** (% by volume): the percent alcohol content of the wine

### Output variable (based on sensory data):
1. **quality** (score between 0 to 10)


## Utilization and Installation of Libraries/Packages
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all the package libraries
# Uncomment the following lines in case the below libraries are not installed on your machine

# install.packages('zipcode', dependencies = T)
# install.packages('dplyr', dependencies = T)
# install.packages('wordcloud', dependencies = T)
# install.packages('grid', dependencies = T)
# install.packages('gridExtra', dependencies = T)
# install.packages('GGally', dependencies = T)
# install.packages('scales', dependencies = T)
# install.packages('RColorBrewer', dependencies = T)
# install.packages('memisc', dependencies = T)
# install.packages('reshape2', dependencies = T)
# install.packages('ggplot2', dependencies = T) 
# install.packages('knitr', dependencies = T)
# install.packages('pastecs', dependencies = T)
# install.packages('corrplot', dependencies = T)
# install.packages('psych', dependencies = T)

library(psych)
library(pastecs)
library(corrplot)
library(ggplot2)
library(Hmisc)
library(dplyr)
library(zipcode)
library(RColorBrewer)
library(wordcloud)
library(grid)
library(gridExtra)
library(GGally)
library(scales)
library(memisc)
library(reshape2)
```

# Descriptive Statistics on different variables

```{r echo=FALSE}
# Load the Data
getwd()
wine <- read.csv('wineQualityWhites.csv')

# Drop unnecessary data
wine$X <- NULL

# List out the variable
names(wine)

# Dimensions Of Dataset
dim(wine) 

# Display the structure of dataset
str(wine)

# Descriptive Statistics
# mean,median,25th and 75th quartiles,min,max
summary(wine)

# Tukey min,lower-hinge, median,upper-hinge,max
fivenum(wine)

# n, nmiss, unique, mean, 5,10,25,50,75,90,95th percentiles 
# 5 lowest and 5 highest scores
describe(wine)
```


# Univariate Plots Section


```{r echo=FALSE, Univariate_Plots}

# Plot each of the input variable
qplot(data = wine, x = fixed.acidity,y = ..density.., geom = "histogram", fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = 'fixed acidity (tartaric acid - g/dm^3)', ylab = "count", binwidth = 1/5)+
  stat_density(geom = "line")

qplot(data = wine, x = volatile.acidity, y = ..density.., geom = "histogram", fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = "volatile acidity (acetic acid - g / dm^3)", ylab = "count", binwidth = 1/100) + stat_density(geom = "line") + stat_density(geom = "line")

qplot(data = wine, x = citric.acid, y = ..density.., geom = "histogram", fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = "citric acid (g / dm^3)", ylab = "count", binwidth = 1/60) + stat_density(geom = "line")

qplot(data = wine, x = residual.sugar, y = ..density.., geom = "histogram", fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = "residual sugar (g / dm^3)", ylab = "count") + stat_density(geom = "line")

qplot(data = wine, x = chlorides, y = ..density.., geom = "histogram", fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = "chlorides (sodium chloride - g / dm^3)", ylab = "count", binwidth = 1/400) + stat_density(geom = "line")

qplot(data = wine, x = free.sulfur.dioxide, y = ..density.., geom = "histogram", fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = "free sulfur dioxide (mg / dm^3)", ylab = "count", binwidth = 3.5) + stat_density(geom = "line")

qplot(data = wine, x = total.sulfur.dioxide, y = ..density.., geom = "histogram", fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = "total sulfur dioxide (mg/dm^3)", ylab = "count", binwidth = 8) + stat_density(geom = "line")

qplot(data = wine, x = density, y = ..density.., geom = "histogram", fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = "density (g / cm^3)", ylab = "count", binwidth = 1/1500) + stat_density(geom = "line")

qplot(data = wine, x = pH, y = ..density.., geom = "histogram", fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = "pH (0 (very acidic) to 14 (very basic))", ylab = "count", binwidth = 1/50) + stat_density(geom = "line")

qplot(data = wine, x = sulphates, y = ..density.., geom = "histogram", fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = "sulphates (potassium sulphate - g / dm3)", ylab = "count", binwidth = 1/100) + stat_density(geom = "line")

qplot(data = wine, x = sulphates, y = ..density.., geom = "histogram", 
      fill=I("blue"), col=I("red"),  alpha=I(.2), 
      xlab = "sulphates (potassium sulphate - g / dm3)", ylab = "count", 
      binwidth = 1/100) +
  stat_density(geom = "line")

# Plot output variable
qplot(data = wine, x = factor(quality), fill=I("blue"), col=I("red"),  alpha=I(.2), xlab = "quality (score between 0 to 10)")

# Numbers of output in each quality
table(wine$quality)

# Categorize wine quality to below_standard, standard, professional in new group
wine$quality.group <- cut(wine$quality, c(0,5,7,10), labels=c("below_standard", "standard", "professional"))

quality.group.plot<-qplot(x = quality.group,data = wine)
grid.arrange(quality.group.plot)

# Numbers of each wine quality
summary(wine$quality.group)
```
#### Understand the distribution
1. Most of the distribtion were skewed to right
2. The peak of volatile acidity is at the point around 0.29
3. The alcohol content is around 8-14
4. Quality has been categorize in three group, most of our white wine is in standard level.

# Univariate Analysis

### What is the structure of your dataset?
This dataset contains 4,898 rows of white wines records data and 13 variables with 11 columns of variables on quantifying the chemical properties of each wine.

### What is/are the main feature(s) of interest in your dataset?
The main interest in this dataset is to identifyt he chemical properties that influence the quality of the white wine. To estimate which chemical brings the best quality to white wines. I was wondering the percentage of alcohol does any impact to the quality of white wines. 

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
The output of 11 variable of chemical properties which influence the quality of white wine.


### Did you create any new variables from existing variables in the dataset?
Yes, we created a new group with cut() that categorize quality of white wine into three standard which is below_standard, standard, professional.
below_standard => 0 ~ 5
standard => 6 ~ 7
professional => 8 ~ 10

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
According to the plot at above, most of the plot were skewed right. The outliers were disappear in the plot chart, I try to get a better quality of chart with modified the bin_width. It's getting better, but not much help. However, I think most investigations should be done later on.


# Bivariate Plots Section


Simple Scatterplot to see the insight of the dataset

```{r echo=FALSE}
attach(wine)
plot(wine$fixed.acidity)
plot(wine$volatile.acidity)
plot(wine$citric.acid)
plot(wine$residual.sugar)
plot(wine$chlorides)
plot(wine$free.sulfur.dioxide)
plot(wine$total.sulfur.dioxide)
plot(wine$density)
plot(wine$pH)
plot(wine$sulphates)
plot(wine$alcohol)
```


```{r echo=FALSE}

p1 <- ggplot(wine, aes(x=quality, y=fixed.acidity)) +
  geom_jitter(alpha = .05)
p2 <- ggplot(wine, aes(x=quality, y=volatile.acidity)) +
  geom_jitter(alpha = .05)
p3 <- ggplot(wine, aes(x=quality, y=citric.acid)) +
  geom_jitter(alpha = .05)
p4 <- ggplot(wine, aes(x=quality, y=residual.sugar)) +
  geom_jitter(alpha = .05)
p5 <- ggplot(wine, aes(x=quality, y=chlorides)) +
  geom_jitter(alpha = .05)
p6 <- ggplot(wine, aes(x=quality, y=free.sulfur.dioxide)) +
  geom_jitter(alpha = .05)
p7 <- ggplot(wine, aes(x=quality, y=total.sulfur.dioxide)) +
  geom_jitter(alpha = .05)
p8 <- ggplot(wine, aes(x=quality, y=density)) +
  geom_jitter(alpha = .05)
p9 <- ggplot(wine, aes(x=quality, y=pH)) +
  geom_jitter(alpha = .05)
p10 <- ggplot(wine, aes(x=quality, y=sulphates)) +
  geom_jitter(alpha = .05)
p11 <- ggplot(wine, aes(x=quality, y=alcohol)) +
  geom_jitter(alpha = .05)

grid.arrange(p1, p2, p3, p4, ncol=2)
grid.arrange(p5, p6, p7, p8, ncol=2)
grid.arrange(p9, p10, p11, ncol=2)
```


```{r echo=FALSE, Bivariate_Plots}

wine.subset <- cor(wine[0:11])
corrplot(
  wine.subset,
  method = "color",
  type = "lower",
  number.cex = 0.60,
  order = "hclust",
  addCoef.col = "black",
  tl.col = "black",
  diag = FALSE # hide correlation coefficient on the principal diagonal
  )

pairs.panels(wine)

```

According to the plot above, We probably can ignore the others correlation due to we found out the correlations.

```{r echo=FALSE}
ggplot(wine, aes(x=residual.sugar, y=alcohol)) +
  geom_jitter(alpha = .05)
ggplot(wine, aes(x=residual.sugar, y=density)) +
  geom_jitter(alpha = .05)
ggplot(wine, aes(x=alcohol, y=density)) +
  geom_jitter(alpha = .05)


```

Here's some insight of residual.sugar, alcohol and density. Seem like the graph have some outlier, which making the plot difficult to observe. I remove the outlier and name is as droped_outlier and visualize it. 


Now, you can observe the graph easily!
```{r echo=FALSE}
droped_outlier <-  wine[wine$residual.sugar < 30,]
ggplot(droped_outlier, aes(x=residual.sugar, y=alcohol, colour = factor(quality))) +
  geom_jitter(alpha = .25)
ggplot(droped_outlier, aes(x=residual.sugar, y=density, colour = factor(quality))) +
  geom_jitter(alpha = .25)
ggplot(droped_outlier, aes(x=alcohol, y=density, colour = factor(quality))) +
  geom_jitter(alpha = .25)


```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
An overview, the alcohol and residual sugar value appears to influences the wine quality. Some other's input variable include free sulfer dioxide, total sulfer dioxide also and density. 

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?
We used the scatterplot to visualize the alcohol/sugar content and density. As you can see, there's some extreme values there!

### What was the strongest relationship you found?
The strongest relationship I found according to the plot is residual sugar with density, then come to the second strongest is alcohol with density.

# Multivariate Plots Section
I'm used the droped_outlier again.
```{r echo=FALSE, Multivariate_Plots}
ggplot(data = droped_outlier, aes(y = residual.sugar, x = alcohol)) +
  geom_point(aes(size = density,color = density)) +
  scale_colour_gradientn(colours=c("red","magenta","blue"))

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?Were there any interesting or surprising interactions between features?
The interesting thing I found at the plot is it's pretty obvious to see the density is getting small when the alcohol is high with low residual sugar.



------

# Final Plots and Summary
### Plot One
```{r echo=FALSE, Plot_One}
qplot(data = wine, x = factor(quality), fill = quality) +
  ggtitle("White Wine Ratings") +
  xlab("Quality (below_standard to professional)") +
  ylab("Counts")
```

### Description One
According to our analysis, most of the white wines rating is in between of 5 to 7. However, they are only fews record on quality with 3, 4, 8 and 9. There's some null record on this dataset like quality with 1, 2 and 10.


### Plot Two
```{r echo=FALSE, Plot_Tw}
ggplot(wine, aes(x = free.sulfur.dioxide, y = alcohol, colour = factor(quality))) + 
  geom_point(size=2.5)
```

### Description Two
The above plot indicates that for the  alcohol and free sulfur dioxide content. The quality of white wines is quite low while the free sulfur dioxide is getting higher. In order to get a high quality white wines, according to this plot seems like we prove again that high alcohol content will influence the quality of white wines. However, the free dioxide should be in the range of 20 to 60 in order to get a good quality.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(data = droped_outlier, aes(y = residual.sugar, x = alcohol)) +
  geom_point(aes(size = density,color = density)) +
  scale_colour_gradientn(colours=c("red","magenta","blue"))
```

### Description Three
While exploring the correlation of alcohol, residual sugar and density, we found some interesting knowledge of white wines with difference chemical properties. The interesting thing I found at the plot is it's pretty obvious to see the density is getting small when the alcohol is high with low residual sugar. If you expecting to have low density, increasing the alcohol content and lowering the residual sugar should be good for you and vice visa.


# Reflection
To get myself familiar with this dataset, I began with extracting the structure of this dataset. This dataset contains 4898 records of white wines and 13 features. Since I don't familiar with this dataset, I try to visualize the dataset from the simplest plot which is histogram. Histogram was doing good to help me get some insight for this dataset. However, to looking for how we get a high quality white wine, I'd visualize some scatterplot to explore the deep insight and get some overview of this. When visualize the dataset, we can't really see it in depth because of outlier, So, we remove it. Hooray! Now, we do one step closer to this dataset. Last but not least, it's always good when the alcohol level is around 12 to 14. If you wish to get a good white wines, this is a simplest thing you should do.

